HO unification from object language to meta language

### 2024-02-27 paper/main.tex
tex add eta example
tex add tricky unif pb
tex

\begin{figure}
  \begin{elpicode}
    type ~(\Eo)~ ftm -> ftm -> o.
    % congruence
    fapp A ~\Eo~fapp B :- forall2 ~(\Eo)~ A B.
    flam F ~\Eo~flam G :- pi x\ x ~\Eo~x => F x ~\Eo~G x.
    fcon C ~\Eo~fcon C.
    fuva N ~\Eo~fuva N.
    % eta
    flam F ~\Eo~T :- pi x\ beta T [x] (T' x),                 ~($\customlabel{clause:eta1}{\eta_1}$)~
      x ~\Eo~x => F x ~\Eo~T' x.
    T ~\Eo~flam F :- pi x\ beta T [x] (T' x),                 ~($\customlabel{clause:eta2}{\eta_2}$)~
      x ~\Eo~x => T' x ~\Eo~F x.
    % beta
    fapp [flam X | TL] ~\Eo~T :- beta (flam X) TL T', T' ~\Eo~T. ~($\customlabel{clause:beta1}{\beta_1}$)~
    T ~\Eo~fapp [flam X | TL] :- beta (flam X) TL T', T ~\Eo~T'. ~($\customlabel{clause:beta2}{\beta_2}$)~
  \end{elpicode}
  \caption{Equal predicate OL}
  \label{code:eq-pred}
  \Description[code:equal-pred]{code:equal-pred}
\end{figure}

\begin{figure}
  \begin{elpicode}
    type ~(\Ee)~ tm -> tm -> o.
    app A ~\Ee~fapp B :- forall2 ~(\Ee)~ A B.
    lam F ~\Ee~flam G :- pi x\ x ~\Ee~x => F x ~\Ee~G x.
    con C ~\Ee~fcon C.
    uva N A ~\Ee~fuva N B :- forall2 ~(\Ee)~ A B.
  \end{elpicode}
  \caption{Equal predicate ML}
  \label{code:eq-pred-meta}
  \Description[code:equal-pred-meta]{code:equal-pred-meta}
\end{figure}

@@

\begin{figure}
\begin{elpicode}
type ~(\Ue)~ tm -> tm -> subst -> subst ->  o.
% Congruence
(app A ~\Ue~app B) S S1 :- fold2 ~\Ue~A B S S1.
(lam F ~\Ue~lam G) S S1 :- 
  pi x\ (pi S\ (x ~\Ue~x) S S) => (F x ~\Ue~G x) S S1.
~\Ue~(con C) (con C) S S.
% deref
(uva N A ~\Ue~T) S S1 :- set N S F,
  beta-abs F A T1, (T1 ~\Ue~T) S S1.
(T ~\Ue~uva N A) S S1 :- set N S F,
  beta-abs F A T1, (T ~\Ue~T1) S S1.
% flex-flex
(uva N A ~\Ue~uva M B) S S3 :-  unset M, unset N,
  pattern-fragment A, pattern-fragment B, 
  new S W S1, prune W Args1 B Ass, 
  assign N S1 Ass S2, assign M S2 Ass S3.
% assignment
(uva N A ~\Ue~T) S S1 :- pattern-fragment A,
  not (T = uva _ _), not_occ N S T,  
  bind A T T1, assign N S T1 S1.
(T ~\Ue~uva N A) S S1 :- pattern-fragment A,
  not (T = uva _ _), not_occ N S T, 
    bind A T T1, assign N S T1 S1.
%  Note: (a infix b) c d === (infix) a b c d

>>>>

\begin{figure}
  \begin{elpicode}
    ~\PYG{k+kd}{type} \PYG{n+nf}{(\Eo)} \PYG{k+kt}{ftm -> ftm -> o}~.
    % congruence
    fapp A ~\Eo~fapp B :- forall2 ~(\Eo)~ A B.
    flam F ~\Eo~flam G :- pi x\ x ~\Eo~x => F x ~\Eo~G x.
    fcon C ~\Eo~fcon C.
    fuva N ~\Eo~fuva N.
    % eta
    flam F ~\Eo~T :- pi x\ beta T [x] (T' x),                 ~($\customlabel{clause:eta1}{\eta_1}$)~
      x ~\Eo~x => F x ~\Eo~T' x.
    T ~\Eo~flam F :- pi x\ beta T [x] (T' x),                 ~($\customlabel{clause:eta2}{\eta_2}$)~
      x ~\Eo~x => T' x ~\Eo~F x.
    % beta
    fapp [flam X | TL] ~\Eo~T :- beta (flam X) TL T', T' ~\Eo~T. ~($\customlabel{clause:beta1}{\beta_1}$)~
    T ~\Eo~fapp [flam X | TL] :- beta (flam X) TL T', T ~\Eo~T'. ~($\customlabel{clause:beta2}{\beta_2}$)~
  \end{elpicode}
  \caption{Equal predicate OL}
  \label{code:eq-pred}
  \Description[code:equal-pred]{code:equal-pred}
\end{figure}

\begin{figure}
  \begin{elpicode}
    ~\PYG{k+kd}{type} \PYG{n+nf}{(\Ee)} \PYG{k+kt}{tm -> tm -> o}~.
    app A ~\Ee~fapp B :- forall2 ~(\Ee)~ A B.
    lam F ~\Ee~flam G :- pi x\ x ~\Ee~x => F x ~\Ee~G x.
    con C ~\Ee~fcon C.
    uva N A ~\Ee~fuva N B :- forall2 ~(\Ee)~ A B.
  \end{elpicode}
  \caption{Equal predicate ML}
  \label{code:eq-pred-meta}
  \Description[code:equal-pred-meta]{code:equal-pred-meta}
\end{figure}

@@

\begin{figure}
\begin{elpicode}
~\PYG{k+kd}{type} \PYG{n+nf}{(\Ue)} \PYG{k+kt}{tm -> tm -> subst -> subst o}~.
% Congruence
(app A ~\Ue~app B) S S1 :- fold2 ~\Ue~A B S S1.
(lam F ~\Ue~lam G) S S1 :- 
  pi x\ (pi S\ (x ~\Ue~x) S S) => (F x ~\Ue~G x) S S1.
~\Ue~(con C) (con C) S S.
% deref
(uva N A ~\Ue~T) S S1 :- set N S F,
  beta-abs F A T1, (T1 ~\Ue~T) S S1.
(T ~\Ue~uva N A) S S1 :- set N S F,
  beta-abs F A T1, (T ~\Ue~T1) S S1.
% flex-flex
(uva N A ~\Ue~uva M B) S S3 :-  unset M, unset N,
  pattern-fragment A, pattern-fragment B, 
  new S W S1, prune W Args1 B Ass, 
  assign N S1 Ass S2, assign M S2 Ass S3.
% assignment
(uva N A ~\Ue~T) S S1 :- pattern-fragment A,
  not (T = uva _ _), not_occ N S T,  
  bind A T T1, assign N S T1 S1.
(T ~\Ue~uva N A) S S1 :- pattern-fragment A,
  not (T = uva _ _), not_occ N S T, 
    bind A T T1, assign N S T1 S1.
%  Note: (a infix b) c d === (infix) a b c d
